---
title: "GangEpl_Analysis_1104"
author: "Jimmy Zhang"
date: "11/4/2020"
output: html_document
---

```{r load packages, cache=TRUE}
library(psych)  # for scatterplot matrix
library(here)  # makes reading data more consistent
library(tidyverse)  # for data manipulation and plotting
library(car)  # some useful functions for regression
library(modelsummary)  # for making tables
library(sjPlot)  # for plotting slopes
library(interactions)  # for plotting interactions
library(haven)  # for importing SPSS/SAS/Stata data
library(lme4)  # for multilevel analysis
library(lattice)  # for dotplot (working with lme4)
library(MuMIn)  # for computing r-squared
library(broom.mixed)  # for summarizing results
library(funModeling)
library(Hmisc)
library(ggplot2)
library(dplyr)
msummary_mixed <- function(models, coef_map = NULL, add_rows = NULL, ...) {
  if (is.null(coef_map)) {
    if (!"list" %in% class(models)) {
      models <- list(models)
    }
    for (model in models) {
      coef_map <- union(coef_map, tidy(model)$term)
    }
  }
  ranef_index <- grep("^(sd|cor)__", x = coef_map)
  coef_map <- c(coef_map[-ranef_index], coef_map[ranef_index])
  names(coef_map) <- coef_map
  rows <- data.frame(term = c("Fixed Effects", "Random Effects"))
  rows <- cbind(rows, rbind(rep("", length(models)), 
                            rep("", length(models))))
  length_fes <- length(coef_map) - length(ranef_index)
  attr(rows, 'position') <- c(1, (length_fes + 1) * 2)
  modelsummary::msummary(models, coef_map = coef_map, add_rows = rows, ...)
}
theme_set(theme_bw())  # Theme; just my personal preference

```

```{r import data}
GangEpl <- read_sav(here("Fall 2020/MMM Lab/Gang and Employment", "GangEpl.sav"))
GangEpl
```

```{r histograms of selected variables}
hist(GangEpl[5:6]) #Histograms of EmployHrs and WmployWg
hist(GangEpl[7:12]) #Histograms of Del1a to Del1f
hist(GangEpl[13:18]) #Histograms of Del2a to Del2f
hist(GangEpl[19]) #Histograms of Del3
hist(GangEpl[20]) #Histograms of Del4
hist(GangEpl[21:26]) #Histograms of Del5a to Del5f
hist(GangEpl[27:30]) #Histograms of GangInv1 to GangInv4
```

```{r scatter plots matrix}
pairs.panels(GangEpl[5:6],  # not plotting the first column
             ellipses = FALSE)
pairs.panels(GangEpl[7:12],  # not plotting the first column
             ellipses = FALSE)
pairs.panels(GangEpl[13:18],  # not plotting the first column
             ellipses = FALSE)
pairs.panels(GangEpl[21:26],  # not plotting the first column
             ellipses = FALSE)
pairs.panels(GangEpl[27:30],  # not plotting the first column
             ellipses = FALSE)
```

```{r scale scores}
average_Del1 <- GangEpl %>% 
  group_by(ID) %>% 
  summarise(average = mean(Del1a + Del1b + Del1c + Del1d + Del1e + Del1f))

average_Del2 <- GangEpl %>% 
  group_by(ID) %>% 
  summarise(average = mean(Del2a + Del2b + Del2c + Del2d + Del2e + Del2f))

average_Del5 <- GangEpl %>% 
  group_by(ID) %>% 
  summarise(average = mean(Del5a + Del5b + Del5c + Del5d + Del5e + Del5f))

average_GangInv <- GangEpl %>% 
  group_by(ID) %>% 
  summarise(average = mean(GangInv1 + GangInv2 + GangInv3 + GangInv4))

average_GangMem <- GangEpl %>% 
  group_by(ID) %>% 
  summarise(average = mean(GangMem1 + GangMem2 + GangMem3 + GangMem4 + GangMem5 + GangMem6 + GangMem7 + GangMem8 + GangMem9 + GangMem10 + GangMem11 + GangMem12 + GangMem13 + GangMem14 + GangMem15)) #Missing data, computing error

average_DC <- GangEpl %>% 
  group_by(ID) %>% 
  summarise(average = mean(DC1 + DC2 + DC3 + DC4 + DC5 + DC6 + DC7 + DC8 + DC9 + DC10 + DC11 + DC12))
```

#An overview of the study design and variables in the GangEpl (Gang Employment) dataset:\

  This is a long form dataset in which each participant was measured multiple times (indicated by the variable Session). We have 337 observations and 57 variables in total.
  
  Within-subject design: Youth were randomly assigned to either BEP or usual services (US), and data were collected at each intervention session. Hypothesis: Increases in employment among BEP participants would be related to reductions in gang involvement over the course of treatment.
  
  Between-subject design: Compare BEP and US participants at the 3 month and 6 month assessment periods. Hypothesis: BEP would lead to higher employment and less gang involvement than US, and that changes in employment would mediate the effect of treatment on gang involvement.
  
  Variables:
  
  ID -> Pariticipant ID
  Age -> The age of participant at each session
  Session -> Time point at which each participant was recorded
  Date -> The date is in year-month-date format (read in haven package)
  EmployHrs -> The hours of working in the current week
  EmployWg -> Hourly wage in the current week
  Del -> Participants' report on the number of gang activities such as stealing over the previous week
  GangInv1 to GangInv4 -> Participants' report on questions about the level of Gang Involvment
  GangMem1 to GangMem15 -> Participants' report on Gang Membership Inventory (the measure of gang membership)
  DC1 to DC12 -> Dummy coded variables for each participant (In the draft paper, the author stated that to identify sources of contamination expected from pooling data from multiple participants, a least squares with dummy variables (LSDV) approach was used. However, I really think a codebook is needed to address how these variables are dummy coded).
  
Concerns: some '88' and '99' value under the column of Del2d and Del2e. Still not really sure what they meant. They are separate from NA.

#An Overview of Data Visualization (Spaghettie Plots and Individual Trajectories) and Scale Scores Computation

```{r cache=TRUE}
GangEpl %>% 
    select(ID, Session, Date, EmployHrs, EmployWg) %>% 
    psych::pairs.panels(jiggle = TRUE, factor = 0.5, ellipses = FALSE, 
                        cex.cor = 1, cex = 0.5)
```
```{r spaghetti plot Part 1}
pEmployHrs <- ggplot(GangEpl, aes(x = Date, y = EmployHrs)) + 
  geom_point() + 
  geom_line(aes(group = ID, color = factor(ID)))   # add lines to connect the data for each person
  # add a mean trajectory
pEmployHrs #The time variable is Date

# p2 <- ggplot(GangEpl, aes(x = Session, y = EmployHrs)) + 
  # geom_point() + 
  # geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  # stat_summary(fun = "mean", col = "red", size = 1, geom = "line")
# p2 #The time variable is Session

pEmployWg <- ggplot(GangEpl, aes(x = Date, y = EmployWg)) + 
  geom_point() + 
  geom_line(aes(group = ID, color = factor(ID)))  # add lines to connect the data for each person
  # add a mean trajectory
pEmployWg #The time variable is Date
```

#Scale Scores

Compute the average scale scores for each scale (Del1, Del2, Del5, GangInv, GangMem):

```{r Computing Scale Score, cache=TRUE}
GangEpl <- GangEpl %>% 
  mutate(Del1 = (Del1a + Del1b + Del1c + Del1d + Del1e + Del1f) / 6, 
         Del2 = (Del2a + Del2b + Del2c + Del2d + Del2e + Del2f) / 6,
         Del5 = (Del5a + Del5b + Del5c + Del5d + Del5e + Del5f) / 6,
         GangInv = (GangInv1 + GangInv2 + GangInv3 + GangInv4) / 4,
         GangMem = (GangMem1 + GangMem2 + GangMem3 + GangMem4 + GangMem5 + GangMem6 + GangMem7 + GangMem8 + GangMem9 + GangMem10 + GangMem11 + GangMem12 + GangMem13 + GangMem14 + GangMem5) / 15)
```

```{r spaghetti plot Part 1}
pDel1 <- ggplot(GangEpl, aes(x = Date, y = Del1)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pDel1

pDel2 <- ggplot(GangEpl, aes(x = Date, y = Del2)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pDel2

pDel3 <- ggplot(GangEpl, aes(x = Date, y = Del3)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pDel3

pDel4 <- ggplot(GangEpl, aes(x = Date, y = Del4)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pDel4

pDel5 <- ggplot(GangEpl, aes(x = Date, y = Del5)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pDel5

pGangInv <- ggplot(GangEpl, aes(x = Date, y = GangInv)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pGangInv

pGangMem <- ggplot(GangEpl, aes(x = Date, y = GangMem)) + 
  geom_point() + 
  geom_line(aes(group = ID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  geom_line(aes(group = ID, color = factor(ID)))
pGangMem
```
Individual Trajectories of EmployHrs and EmployWg:

```{r Individual Trajectories}
GangEpl %>% 
  ggplot(aes(x = Date)) + 
  geom_line(aes(y = EmployHrs), col = "blue") + 
  geom_line(aes(y = EmployWg), col = "red") + 
  facet_wrap( ~ ID, ncol = 4)
```

```{r within-person and between-person associations}
GangEpl <- GangEpl %>% 
  group_by(ID) %>% 
  mutate(across(c(EmployHrs, EmployWg), 
                # The `.` means the variable to be operated on
                list("pm" = ~ mean(., na.rm = TRUE), 
                     "pmc" = ~ . - mean(., na.rm = TRUE)))) %>% 
  ungroup()

pAssociations <- GangEpl %>% 
  ggplot(aes(x = EmployHrs, y = EmployWg, col = ID)) + 
  geom_point(size = 0.7, alpha = 0.5) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(x = EmployHrs_pm, y = EmployWg_pm, fill = ID), 
             shape = 24, col = "black", size = 2) + 
  geom_smooth(aes(x = EmployHrs_pm, y = EmployWg_pm, group = 1), 
              method = "lm", linetype = "dashed", fullrange = TRUE)
pAssociations
```

#Random Intercept Model

I ran the intercept models for each variable for possible future analysis.

```{r random Intercep Model of each variable}
ran_int_EmployHrs <- lmer(EmployHrs ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_EmployHrs)

ran_int_EmployWg <- lmer(EmployWg ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_EmployWg)

ran_int_Del1 <- lmer(Del1 ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_Del1)

ran_int_Del2 <- lmer(Del2 ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_Del2)

ran_int_Del3 <- lmer(Del3 ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_Del3)

ran_int_Del4 <- lmer(Del4 ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_Del4)

ran_int_Del5 <- lmer(Del5 ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_Del5)

ran_int_GangInv <- lmer(GangInv ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_GangInv)

ran_int_GangMem <- lmer(GangMem ~ 1 + (1 | ID), data = GangEpl)
summary(ran_int_GangMem)
```

#Adding variables

```{r}
ModelGangInv <- lmer(GangInv ~ EmployHrs + (EmployHrs | ID), data = GangEpl)
summary(ModelGangInv)
```

